Insights of Even-point Multi-loop Unitarity: Exponentiation, Anomalies and Evanescence

### 2023-06-13 eventPointTwoLoop.tex
Section 2 and section 3
added to background on EP theories, added full write up of Unitarity methods, D-dim and 4D, background on U(1) HD correctios in abelian OSS, defined precisely what the field means by a recursive integral (plot twist, no one states it explicitly) -- drew a bunch of new diagrams for recursive integrals and Forde's integral reduction. Added 4D vs D-dimensional formal polarizations and explained our procedure for deriving all the nice tensor structures

\newcommand{\scaleIntCscalar}[4]{ {
\begin{tikzpicture}[baseline=(current  bounding  box.center)]
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {#1};
\vertex (a2) at (-.8, 0.6) {#2};
\vertex [dot,scale=2](mid1) at (0,0){};
\vertex [dot,scale=1.5,hgrey0](mid2) at (0,0){};
\vertex [dot,scale=2](mid3) at (1,0){};
\vertex [dot,scale=1.5,hgrey0](mid4) at (1,0){};
\vertex [dot,scale=2](mid5) at (2,0){};
\vertex [dot,scale=1.5,hgrey0](mid6) at (2,0){};
\vertex (a3) at (2.8, .6) {#3};
\vertex (a4) at (2.8, -.6) {#4};
\diagram{
(mid1) --[ ultra thick,](a1),
(mid1) --[ ultra thick,](a2),
(mid1) --[ ultra thick,out=60,in=120,min distance=0.4cm](mid3),
(mid1) --[ ultra thick,out=-60,in=-120,min distance=0.4cm](mid3),
(mid3) --[ ultra thick,out=60,in=120,min distance=0.4cm](mid5),
(mid3) --[ ultra thick,out=-60,in=-120,min distance=0.4cm](mid5),
(mid5) --[ ultra thick](a4),
(mid5) --[ ultra thick,](a3)
};
\end{feynman}
\end{tikzpicture}
}
}

@@

@@

\abstract{
We identify novel  structure in newly computed multi-loop amplitudes and quantum actions for even-point effective field theories, including both the nonlinear sigma model (NLSM) and double-copy gauge theories such as Born-Infeld and its supersymmetric generalizations.  We exploit special properties of these even-point theories to generalize the direct extraction of Feynman integral coefficients using $D$-dimensional unitarity cuts to the multi-loop case for the first time. In doing so, we find that the leading IR divergence of NLSM amplitudes exponentiates. We then systematically compute the two-loop anomalous behavior of Born-Infeld, and find that the counterterms required to restore $U(1)$ behavior at loop-level in the Born-Infeld theory can be constructed via a symmetric-structure double-copy.  We also demonstrate that the one-minus $(-+++)$ two-loop amplitude vanishes upon the introduction of an evanescent operator in the $D$-dimensional description of Born-Infeld theory. In addition to these pure photon counterterms, we verify through explicit calculation that the anomalous matrix elements that violates $U(1)$ duality invariance can be alternatively cancelled by summing over internal $\mathcal{N}=4$ DBIVA superfields. Finally we find that $\mathcal{N}=4$ Dirac-Born-Infeld-Volkov-Akulov (DBIVA) amplitudes permit double-copy construction through two-loop order. This is shown by reproducing our unitarity based result with a loop-level double copy between $\mathcal{N}=4$ super-Yang-Mills numerators and our two-loop NLSM amplitudes. This result provides strong evidence for the existence of a color-dual representations for NLSM beyond one-loop. We conclude with an overview of how $D$-dimensional four-photon counterterms can be constructed in generality with the symmetric-structure double-copy, and outline a convenient way of counting evanescent operators using Hilbert series as generating functions. }

@@

For gauge theories, it is often useful to further decompose the full color-dressed amplitudes into kinematic building blocks where the color information is stripped away. For example, at tree level, gauge theories $n$-point amplitudes can be re-expressed in terms of a trace basis decomposition as follows:
\be
\mathcal{A}_{n,\text{tree}} = \sum_{\sigma \in S^{n-1}} \text{Tr}(T^1 T^{\sigma(2)}\cdots T^{\sigma(n)}) A(1,\sigma(2),...,\sigma(n))
\ee
where the sum is taken over the inequivalent orderings of group theory generator traces. A similar decomposition applies at one-loop gauge theory amplitudes, which we will describe in \sect{sec:Loops}. The color-ordered functions are called \textit{partial amplitudes}, as they only contain on-shell information for a particular color ordering. However, like full color-dressed amplitudes, they factor on poles to lower-point partial amplitudes, and when describe vector theories they are independently gauge invariant since they weight linearly independent color traces.

@@

When working with kinematics in arbitrary dimesions, we will use formal Lorentz covariant polarizations, $\varepsilon^\mu_a$, and momenta, $k^\mu_a$. The mostly minus signature will be used throughout. The only restrictions we will place on the momenta and polarizations are the standard on-shell constraints, momentum conservation and the null-momenta condition,
\be
\sum_{a} k_a^\mu =0 \qquad \qquad k_a^2 =0
\ee
Furthermore, we will take external polarizations to be in a transverse gauge throughout:
\be
\varepsilon_a \cdot k_a = 0
\ee
Formal polarizations will always be expressed as $\varepsilon_a^\mu$ to distinguish between the extra dimensional bit of the integration measure in $D=4-2\epsilon$. As noted, it will be convenient to take $D\rightarrow 4$ in certain circumstances after integration. We will use the conventions of ref. \cite{jjmcTASI2014} throughout the text, which we provide below
\be
s_{ab} = \langle ab \rangle[ba]
\ee
where $s_{ab} = (k_a+k_b)^2$. Furthermore, polarization dot products with particular helicity states can be mapped as follows:
\begin{align}
k_a \cdot \varepsilon_b^{(+)} &= \frac{\langle q a \rangle[ab]}{\sqrt{2}\langle q b\rangle}
\qquad\quad \qquad
k_a \cdot \varepsilon_b^{(-)} = -\frac{[qa]\langle ab\rangle}{\sqrt{2}[qb]}
\\
\varepsilon_a^{(-)}\cdot \varepsilon_b^{(+)} &= - \frac{\langle q a\rangle [qb]}{ [qa]\langle q b\rangle} 
\qquad \qquad
\varepsilon_a^{(\pm)}\cdot \varepsilon_b^{(\pm)} = 0 
\end{align}
where $q^\mu$ is some null reference momentum that all polarizations are projected along. In practice, rather than plugging in these 4D expressions by hand and applying a series of Feirz Identities \textit{ad nauseum}, the results obtained throughout the text are constructed by numerically evaluating the dot products and projecting them along spanning set of helicity configurations. For easy access, below we provide the component definition of our spinor bracket that are consistent with the conventions above,
\begin{align}
\langle ab \rangle &= \frac{(a_1 + i a_2)(b_0+b_3)-(b_1 + i b_2)(a_0+a_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}
\\
[ab] &= \frac{(b_1 - i b_2)(a_0+a_3)-(a_1 - i a_2)(b_0+b_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}
\end{align}
where the $a_i$ are coordinate values of the four-vector, $k^\mu_a = (a_0,a_1,a_2,a_3)$. With this in hand, we will transition away from on-shell methods and venture into the Lagrangian descriptions of the theories that we will study throughout the text. 
\subsection{Even-point Effective Field Theories}
\label{subsec:EPEFTReview}
\dj{Some background here about symmetries, and constraininng field theories, yada yada yada}
\paragraph{Simplicity of Even-point EFT}

By far the most important feature of these amplitudes relevant for our analysis is that they are generated by purely \textit{even-point} (EP) contacts. That is, the effective actions of EP theories are consistent with the following interaction Lagrangian:
\begin{equation}
\mathcal{L}^{\text{EP}}_{\text{int.}} = \sum_{k=2} \alpha'^{n-2} \mathcal{O}^{(2n)}_{2k} 
\end{equation}
where $2k$ counts the field content of the effective operator, and $2n$ counts the mass-dimension determined by the number of fields and derivatives contained in the operator. The dimensionful scale, $\alpha'$, carries mass dimension $[\alpha']=-2$ in $D=4$.

The even-point nature of this class of theories make them comparatively easy to compute relative to their arbitrary-multiplicity cousins. Indeed, the observation that will drive many of the multi-loop results is the following: \textit{amplitudes for EP theories at next-to-next-to-leading-order (NNLO) in $\alpha'$ are recursively one-loop}. To gain some familiarity with why this is the case, consider the interactions needed to construct two loop amplitudes for a simple $\varphi^{2k}$-scalar theory:
\begin{equation}\label{eq:evenPointL}
\mathcal{L}_{\varphi^{2k}} = \frac{1}{2}(\partial \varphi)^2 + c_4 \varphi^4 + c_6 \varphi^6+ c_8 \varphi^8+\cdots 
\end{equation}
While the Feynman rules for this theory are trivial, unlike those needed for the vectors and fermions in the DBIVA supermultiplet and the derivatively coupled pions in NLSM, the Feynman diagrams needed to compute amplitudes are identical to those of $A^{\text{OSS}}_{\times}$. Indeed, the two-loop amplitude for $\varphi^{2k}$ theory is generated simply by evaluating the following set of scalar 1-particle-reducible (1PI) integrals:

@@

Since the external states are massless, only the first two integrals are non-vanishing in dimensional regularization -- the rest are scaleless. Therefore, at least perturbatively, the above Even-point theory of \eqn{eq:evenPointL} is equivalent to $\varphi^4$ theory through two-loop order: 
\begin{equation}
\mathcal{M}_{\varphi^{2k}}^{\text{2-loop}} \equiv \mathcal{M}_{\varphi^{4}}^{\text{2-loop}} 
\end{equation}
In addition to being perturbatively equivalent to the quartic sector of the theory at two-loop, we can see above that the integrals for the Feynman diagrams in \eqn{eq:2loopEven} are \textit{recursively one-loop}. In other words, to integrate the full amplitude at two-loop, we only need an iterated basis of one-loop integrals. This observation comes with the added advantage of permitting the application of one-loop Passarino-Veltman tensor reduction \cite{Passarino:1978jh} to two loop integrals. This property of recursively one-loop integrals will dramatically simplify the higher derivative EP theories that we consider in this work.

In what follows, we provide a catalog of the even-point effective field theories studied in this paper. We will define the Lagrangians  be expressed in a spacetime coordinate frame, such that, $\mathcal{L}$, is the integrand for a specific theory, to be the integrand of the action, $S$, when working in spacetime coordinate frame:
\be
S= \int d^D x \,\mathcal{L}
\ee

>>>>

\newcommand{\scaleIntBubbleProbeA}[2]{ {
\begin{tikzpicture}[baseline=-2.5]
\begin{feynman}
\vertex (a1) at (-.6,.3) {#1};
\vertex (a2) at (-.6, -.3) {#2};
\vertex (a4) at (.5, .7) {${}^{\alpha_1=1}$};
\vertex (a6) at (.5, -.8) {${}^{\alpha_4=1}$};
\vertex [dot,scale=2](mid1) at (0,0){};
\vertex [dot,scale=1.5,hgrey0](mid2) at (0,0){};
\vertex [dot,scale=2](mid5) at (1,0){};
\vertex [dot,scale=1.5,hgrey0](mid6) at (1,0){};
\vertex (a3) at (1.6, .3) {#2};
\vertex (a4) at (1.6, -.3) {#2};
\diagram{
(a1) --[ultra thick,](mid1),
(a2) --[ultra thick,](mid1),
(mid1) --[ ultra thick,out=60,in=120,min distance=0.5cm](mid5),
(mid1) --[ ultra thick,out=30,in=150,min distance=0.2cm](mid5),
(mid1) --[ ultra thick,out=-30,in=-150,min distance=0.2cm](mid5),
(mid1) --[ ultra thick,out=-60,in=-120,min distance=0.5cm](mid5),
(mid5) --[ultra thick,](a3),
(mid5) --[ultra thick,](a4)
};
\end{feynman}
\end{tikzpicture}
}
}

\newcommand{\scaleIntBubbleProbeB}[2]{ {
\begin{tikzpicture}[baseline=-2.5]
\begin{feynman}
\vertex (a1) at (-.6,.3) {#1};
\vertex (a2) at (-.6, -.3) {#2};
\vertex [dot,scale=2](mid1) at (0,0){};
\vertex (a4) at (.5, .5) {${}^{\alpha_2=D/2-2}$};
\vertex (a6) at (.5, -.6) {${}^{\alpha_4=1}$};
\vertex [dot,scale=1.5,hgrey0](mid2) at (0,0){};
\vertex [dot,scale=2](mid5) at (1,0){};
\vertex [dot,scale=1.5,hgrey0](mid6) at (1,0){};
\vertex (a3) at (1.6, .3) {#2};
\vertex (a4) at (1.6, -.3) {#2};
\diagram{
(a1) --[ultra thick,](mid1),
(a2) --[ultra thick,](mid1),
(mid1) --[ ultra thick,out=45,in=135,min distance=0.3cm](mid5),
(mid1) --[ ultra thick](mid5),
(mid1) --[ ultra thick,out=-45,in=-135,min distance=0.3cm](mid5),
(mid5) --[ultra thick,](a3),
(mid5) --[ultra thick,](a4)
};
\end{feynman}
\end{tikzpicture}
}
}

\newcommand{\scaleIntBubbleProbeC}[2]{ {
\begin{tikzpicture}[baseline=-2.5]
\begin{feynman}
\vertex (a1) at (-.6,.3) {#1};
\vertex (a2) at (-.6, -.3) {#2};
\vertex (a4) at (.5, .4) {${}^{\alpha_3=D/2-2}$};
\vertex (a6) at (.5, -.5) {${}^{\alpha_4=1}$};
\vertex [dot,scale=2](mid1) at (0,0){};
\vertex [dot,scale=1.5,hgrey0](mid2) at (0,0){};
\vertex [dot,scale=2](mid5) at (1,0){};
\vertex [dot,scale=1.5,hgrey0](mid6) at (1,0){};
\vertex (a3) at (1.6, .3) {#2};
\vertex (a4) at (1.6, -.3) {#2};
\diagram{
(a1) --[ultra thick,](mid1),
(a2) --[ultra thick,](mid1),
(mid1) --[ ultra thick,out=30,in=150,min distance=0.2cm](mid5),
(mid1) --[ ultra thick,out=-30,in=-150,min distance=0.2cm](mid5),
(mid5) --[ultra thick,](a3),
(mid5) --[ultra thick,](a4)
};
\end{feynman}
\end{tikzpicture}
}
}

\newcommand{\scaleIntBubbleProbe}[2]{ {
\begin{tikzpicture}[baseline=-2]
\begin{feynman}
\vertex (b1) at (-.8,.8) {};
\vertex (b2) at (1.8,.8) {};
\vertex (b3) at (1.8,-.8) {};
\vertex (b4) at (-.8,-.8) {};
\vertex (a1a) at (-.8,.6) {};
\vertex (a1b) at (-.6,.1) {$\vdots$};
\vertex (a1c) at (-.8,-.6) {};
\vertex (a2) at (.6,.7) {$\alpha_1$};
\vertex (a4) at (.6,-.7) {$\alpha_2$};
\vertex [dot,scale=2](mid1) at (0,0){};
\vertex [dot,scale=1.5,hgrey0](mid2) at (0,0){};
\vertex [dot,scale=2](mid5) at (1,0){};
\vertex [dot,scale=1.5,hgrey0](mid6) at (1,0){};
\vertex (a3a) at (1.8, .6) {};
\vertex (a3b) at (1.6, .1) {$\vdots$};
\vertex (a3c) at (1.8,-.6) {};
\diagram{
(b1) --[scalar, hblue,ultra thick,](b4),
(b3) --[scalar, hblue,ultra thick,](b2),
(a1a) --[fermion, ultra thick,](mid1),
(a1c) --[fermion, ultra thick,](mid1),
(mid1) --[ ultra thick,out=60,in=120,min distance=0.4cm](mid5),
(mid1) --[ ultra thick,out=-60,in=-120,min distance=0.4cm](mid5),
(mid5) --[fermion, ultra thick,](a3a),
(mid5) --[fermion, ultra thick,](a3c)
};
\end{feynman}
\end{tikzpicture}
}
}

\newcommand{\scaleIntCscalar}[4]{ {
\begin{tikzpicture}[baseline=(current  bounding  box.center)]
\begin{feynman}
\vertex (a1) at (-.6, -0.6) {#1};
\vertex (a2) at (-.6, 0.6) {#2};
\vertex [dot,scale=2](mid1) at (0,0){};
\vertex [dot,scale=1.5,hgrey0](mid2) at (0,0){};
\vertex [dot,scale=2](mid3) at (.8,0){};
\vertex [dot,scale=1.5,hgrey0](mid4) at (.8,0){};
\vertex [dot,scale=2](mid5) at (1.6,0){};
\vertex [dot,scale=1.5,hgrey0](mid6) at (1.6,0){};
\vertex (a3) at (2.2, .6) {#3};
\vertex (a4) at (2.2, -.6) {#4};
\diagram{
(mid1) --[ ultra thick,](a1),
(mid1) --[ ultra thick,](a2),
(mid1) --[ ultra thick,out=60,in=120,min distance=0.4cm](mid3),
(mid1) --[ ultra thick,out=-60,in=-120,min distance=0.4cm](mid3),
(mid3) --[ ultra thick,out=60,in=120,min distance=0.4cm](mid5),
(mid3) --[ ultra thick,out=-60,in=-120,min distance=0.4cm](mid5),
(mid5) --[ ultra thick](a4),
(mid5) --[ ultra thick,](a3)
};
\end{feynman}
\end{tikzpicture}
}
}

@@

\newcommand{\bubCutStates}[4]{ {
\begin{tikzpicture}[baseline=-4]
\begin{feynman}
\vertex (a1) [dot,scale=1.6]at (-.6,0){};
\vertex (a2) [dot,scale=1.6]at (.6,0){};
\vertex (a1) [dot,hgrey0,scale=1.3]at (-.6,0){};
\vertex (a2) [dot,hgrey0,scale=1.3]at (.6,0){};
\vertex (a1a) at (-1.3,.6){#2};
\vertex (a1b) at (-1.3,-.6){#1};
\vertex (a1c) at (-1.1,0){};
\vertex (a2a) at (1.3,.6){#3};
\vertex (a2b) at (1.3,-.6){#4};
\vertex (a2c) at (1.1,0){};
\vertex (a5) at (0,.4){};
\vertex (a7) at (0,-.4){};
\vertex (a8) at (-.5,0){};
\vertex (b1) at (0,.7);
\vertex (b2) at (0,-.7);
\vertex (b3) at (.7,0);
\vertex (b4) at (-.7,0);
\vertex (b5) at (0,0){};
\diagram{
(b1) --[scalar,hblue,ultra thick,](b5),
(b2) --[scalar,hblue,ultra thick,](b5),
(a5) --[ ultra thick,in=60,out=180,min distance=0.1cm](a1),
(a5) --[ ultra thick,out=0,in=120,min distance=0.1cm](a2),
(a2) --[ ultra thick,out=-120,in=0,min distance=0.1cm](a7),
(a1) --[ ultra thick,out=-60,in=180,min distance=0.1cm](a7),
(a1a) --[ ultra thick,](a1),
(a1b) --[ ultra thick,](a1),
(a2a) --[ ultra thick,](a2),
(a2b) --[ ultra thick,](a2),
};
\end{feynman}
\end{tikzpicture}
}
}

\newcommand{\bubCut}{ {
\begin{tikzpicture}[baseline=-4]
\begin{feynman}
\vertex (a1) [dot,scale=1.6]at (-.6,0){};
\vertex (a2) [dot,scale=1.6]at (.6,0){};
\vertex (a1) [dot,hgrey0,scale=1.3]at (-.6,0){};
\vertex (a2) [dot,hgrey0,scale=1.3]at (.6,0){};
\vertex (a1a) [dot,scale=1.6]at (-.9,.25);
\vertex (a1b) [dot,scale=1.6]at (-.9,-.25);
\vertex (a1c) at (-1.1,-0.05){};
\vertex (a2a) [dot,scale=1.6]at (.9,.25);
\vertex (a2b) [dot,scale=1.6]at (.9,-.25);
\vertex (a2c) at (1.1,-0.05){${}^{P_i}$};
\vertex (a5) at (0,.4){};
\vertex (a7) at (0,-.4){};
\vertex (a8) at (-.5,0){};
\vertex (b1) at (0,.7);
\vertex (b2) at (0,-.7);
\vertex (b3) at (.7,0);
\vertex (b4) at (-.7,0);
\vertex (b5) at (0,0){};
\diagram{
(b1) --[scalar,hblue,ultra thick,](b5),
(b2) --[scalar,hblue,ultra thick,](b5),
(a5) --[ ultra thick,in=60,out=180,min distance=0.1cm](a1),
(a5) --[ ultra thick,out=0,in=120,min distance=0.1cm](a2),
(a2) --[ ultra thick,out=-120,in=0,min distance=0.1cm](a7),
(a1) --[ ultra thick,out=-60,in=180,min distance=0.1cm](a7),
(a1a) --[ ultra thick,](a1),
(a1b) --[ ultra thick,](a1),
(a2a) --[ ultra thick,](a2),
(a2b) --[ ultra thick,](a2),
};
\end{feynman}
\end{tikzpicture}
}
}

\newcommand{\triCut}{ {
\begin{tikzpicture}[baseline=-4]
\begin{feynman}
\vertex (a1) [dot,scale=1.6] at (-.5,.5){};
\vertex (a2) [dot,scale=1.6]at (.6,0){};
\vertex (a4) [dot,scale=1.6]at (-.5,-.5){};
\vertex (a1) [dot,hgrey0,scale=1.3] at (-.5,.5){};
\vertex (a2) [dot,hgrey0,scale=1.3]at (.6,0){};
\vertex (a4) [dot,hgrey0,scale=1.3]at (-.5,-.5){};
\vertex (a1a) [dot,scale=1.6] at (-.5,.8);
\vertex (a1b) [dot,scale=1.6] at (-.8,.5);
\vertex (a1c) at (-.8,.8){${}^{P_i}$};
\vertex (a2a) [dot,scale=1.6]at (.9,.25);
\vertex (a2b) [dot,scale=1.6]at (.9,-.25);
\vertex (a2c) at (1.1,-0.1){${}^{P_j}$};
\vertex (a4a) [dot,scale=1.6]at (-.5,-.8);
\vertex (a4b) [dot,scale=1.6]at (-.8,-.5);
\vertex (a4c) at (-.8,-.9){};
\vertex (a5) at (0,.3){};
\vertex (a7) at (0,-.3){};
\vertex (a8) at (-.5,0){};
\vertex (b1) at (0,.7);
\vertex (b2) at (0,-.7);
\vertex (b3) at (.7,0);
\vertex (b4) at (-.7,0);
\vertex (b5) at (0,0){};
\diagram{
(b1) --[scalar,hblue,ultra thick,](b5),
(b2) --[scalar,hblue,ultra thick,](b5),
(b4) --[scalar,hblue,ultra thick,](b5),
(a5) --[ ultra thick,](a1),
(a5) --[ ultra thick,](a2),
(a2) --[ ultra thick,](a7),
(a4) --[ ultra thick,](a7),
(a8) --[ ultra thick,](a4),
(a1) --[ ultra thick,](a8),
(a1a) --[ ultra thick,](a1),
(a1b) --[ ultra thick,](a1),
(a2a) --[ ultra thick,](a2),
(a2b) --[ ultra thick,](a2),
(a4a) --[ ultra thick,](a4),
(a4b) --[ ultra thick,](a4),
};
\end{feynman}
\end{tikzpicture}
}
}

\newcommand{\boxCut}{ {
\begin{tikzpicture}[baseline=-4]
\begin{feynman}
\vertex (a1) [dot,scale=1.6] at (-.5,.5){};
\vertex (a2) [dot,scale=1.6]at (.5,.5){};
\vertex (a3) [dot,scale=1.6]at (.5,-.5){};
\vertex (a4) [dot,scale=1.6]at (-.5,-.5){};
\vertex (a1) [dot,hgrey0,scale=1.3] at (-.5,.5){};
\vertex (a2) [dot,hgrey0,scale=1.3]at (.5,.5){};
\vertex (a3) [dot,hgrey0,scale=1.3]at (.5,-.5){};
\vertex (a4) [dot,hgrey0,scale=1.3]at (-.5,-.5){};
\vertex (a1a) [dot,scale=1.6] at (-.5,.8);
\vertex (a1b) [dot,scale=1.6] at (-.8,.5);
\vertex (a1c) at (-.8,.8){${}^{P_i}$};
\vertex (a2a) [dot,scale=1.6]at (.5,.8);
\vertex (a2b) [dot,scale=1.6]at (.8,.5);
\vertex (a2c) at (.9,.8){${}^{P_j}$};
\vertex (a3a) [dot,scale=1.6]at (.5,-.8);
\vertex (a3b) [dot,scale=1.6]at (.8,-.5);
\vertex (a3c) at (.9,-.9){${}^{P_k}$};
\vertex (a4a) [dot,scale=1.6]at (-.5,-.8);
\vertex (a4b) [dot,scale=1.6]at (-.8,-.5);
\vertex (a3c) at (-.8,-.9){};
\vertex (a5) at (0,.5){};
\vertex (a6) at (.5,0){};
\vertex (a7) at (0,-.5){};
\vertex (a8) at (-.5,0){};
\vertex (b1) at (0,.7);
\vertex (b2) at (0,-.7);
\vertex (b3) at (.7,0);
\vertex (b4) at (-.7,0);
\vertex (b5) at (0,0){};
\diagram{
(b1) --[scalar,hblue,ultra thick,](b5),
(b2) --[scalar,hblue,ultra thick,](b5),
(b3) --[scalar,hblue,ultra thick,](b5),
(b4) --[scalar,hblue,ultra thick,](b5),
(a5) --[ ultra thick,](a1),
(a5) --[ ultra thick,](a2),
(a2) --[ ultra thick,](a6),
(a6) --[ ultra thick,](a3),
(a4) --[ ultra thick,](a7),
(a7) --[ ultra thick,](a3),
(a8) --[ ultra thick,](a4),
(a1) --[ ultra thick,](a8),
(a1a) --[ ultra thick,](a1),
(a1b) --[ ultra thick,](a1),
(a2a) --[ ultra thick,](a2),
(a2b) --[ ultra thick,](a2),
(a3a) --[ ultra thick,](a3),
(a3b) --[ ultra thick,](a3),
(a4a) --[ ultra thick,](a4),
(a4b) --[ ultra thick,](a4),
};
\end{feynman}
\end{tikzpicture}
}
}

@@

\abstract{
We identify novel  structure in newly computed multi-loop amplitudes and quantum actions for even-point effective field theories, including both the nonlinear sigma model (NLSM) and double-copy gauge theories such as Born-Infeld and its supersymmetric generalizations.  For the first time, we exploit special properties of these even-point theories to generalize the direct extraction of Feynman integral coefficients using $D$-dimensional unitarity cuts to the multi-loop case. In doing so, we find that the leading IR divergence of NLSM amplitudes exponentiates. We then systematically compute the two-loop anomalous behavior of Born-Infeld, and find that the counterterms required to restore $U(1)$ behavior at loop-level in the Born-Infeld theory can be constructed via a symmetric-structure double-copy.  We also demonstrate that the one-minus $(-+++)$ two-loop amplitude vanishes upon the introduction of an evanescent operator in the $D$-dimensional description of Born-Infeld theory. In addition to these pure photon counterterms, we verify through explicit calculation that the anomalous matrix elements that violates $U(1)$ duality invariance can be alternatively cancelled by summing over internal $\mathcal{N}=4$ DBIVA superfields. Finally we find that $\mathcal{N}=4$ Dirac-Born-Infeld-Volkov-Akulov (DBIVA) amplitudes permit double-copy construction through two-loop order. This is shown by reproducing our unitarity based result with a loop-level double copy between $\mathcal{N}=4$ super-Yang-Mills numerators and our two-loop NLSM amplitudes. This result provides strong evidence for the existence of a color-dual representations for NLSM beyond one-loop. We conclude with an overview of how $D$-dimensional four-photon counterterms can be constructed in generality with the symmetric-structure double-copy, and outline a convenient way of counting evanescent operators using Hilbert series as generating functions. }

@@

For gauge theories, it is often useful to further decompose the full color-dressed amplitudes into kinematic building blocks where the color information is stripped away. For example, at tree level, gauge theories $n$-point amplitudes can be re-expressed in terms of a trace basis decomposition as follows:
\be
\mathcal{A}_{n,\text{tree}} = \sum_{\sigma \in S^{n-1}} \text{Tr}(T^1 T^{\sigma(2)}\cdots T^{\sigma(n)}) A_n(1,\sigma(2),...,\sigma(n))
\ee
where the sum is taken over the inequivalent orderings of group theory generator traces. A similar decomposition applies at one-loop gauge theory amplitudes, which we will describe in \sect{sec:Loops}. The color-ordered functions are called \textit{partial amplitudes}, as they only contain on-shell information for a particular color ordering. However, like full color-dressed amplitudes, they factor on poles to lower-point partial amplitudes, and when describe vector theories they are independently gauge invariant since they weight linearly independent color traces.

@@

When working with kinematics in arbitrary dimesions, we will use formal Lorentz covariant polarizations, $\varepsilon^\mu_a$, and momenta, $k^\mu_a$. The mostly minus signature will be used throughout. The only restrictions we will place on the momenta and polarizations are the standard on-shell constraints, momentum conservation and the null-momenta condition,
\sum_{a} k_a^\mu =0 \qquad \qquad k_a^2 =0
Furthermore, we will take external polarizations to be in a transverse gauge throughout:
\varepsilon_a \cdot k_a = 0
Formal polarizations will always be expressed as $\varepsilon_a^\mu$ to distinguish between the extra dimensional bit of the integration measure in $D=4-2\epsilon$. As noted, it will be convenient to take $D\rightarrow 4$ in certain circumstances after integration. We will use the conventions of ref. \cite{jjmcTASI2014} throughout the text, which we provide below
s_{ab} = \langle ab \rangle[ba]
where $s_{ab} = (k_a+k_b)^2$. Furthermore, polarization dot products with particular helicity states can be mapped as follows:
k_a \cdot \varepsilon_b^{(+)} &= \frac{\langle q a \rangle[ab]}{\sqrt{2}\langle q b\rangle}
\qquad\quad \qquad
k_a \cdot \varepsilon_b^{(-)} = -\frac{[qa]\langle ab\rangle}{\sqrt{2}[qb]}
\varepsilon_a^{(-)}\cdot \varepsilon_b^{(+)} &= - \frac{\langle q a\rangle [qb]}{ [qa]\langle q b\rangle} 
\qquad \qquad
\varepsilon_a^{(\pm)}\cdot \varepsilon_b^{(\pm)} = 0 
where $q^\mu$ is some null reference momentum that all polarizations are projected along. In practice, rather than plugging in these 4D expressions by hand and applying a series of Feirz Identities \textit{ad nauseum}, the results obtained throughout the text are constructed by numerically evaluating the dot products and projecting them along spanning set of helicity configurations. For easy access, below we provide the component definition of our spinor bracket that are consistent with the conventions above,
\langle ab \rangle &= \frac{(a_1 + i a_2)(b_0+b_3)-(b_1 + i b_2)(a_0+a_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}
[ab] &= \frac{(b_1 - i b_2)(a_0+a_3)-(a_1 - i a_2)(b_0+b_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}
where the $a_i$ are componet values of the four-vector, $k^\mu_a = (a_0,a_1,a_2,a_3)$. With this in hand, we will transition away from on-shell methods and venture into the Lagrangian descriptions of the theories that we will study throughout the text. 
\subsection{Even-point Effective Field Theories}
\label{subsec:EPEFTReview}
In this section we provide some background on the even-point effective field theories that we study throughout the paper. Typically the power of EFT is in its ability to parameterize hidden UV dynamics in \textit{a priori} unconstrained Wilson coefficients. However, throughout the text we will focus on theories whose Wilson coefficients are dictated by symmetry. The simplicity of theories we choose to study belies the universality of our results, since it turns out that only the \textit{leading order} four-point interactions will contribute to perturbative physics through two-loop order. We demonstrate this below with an explicit example from $\varphi^{2k}$-theory. 
\paragraph{Simplicity of Even-point EFT} As alluded to, the even-point nature of the theories studied in this paper make them comparatively easy to compute relative to their arbitrary-multiplicity cousins. Indeed, the observation that will drive many of the multi-loop results is the following: \textit{amplitudes for EP theories at next-to-next-to-leading-order (NNLO) are recursively one-loop}. To gain some familiarity with why this is the case, consider the interactions needed to construct two loop amplitudes for a simple $\varphi^{2k}$-scalar theory:
\begin{equation}\label{eq:evenPointL}
\mathcal{L}_{\varphi^{2k}} = \frac{1}{2}(\partial \varphi)^2 + c_4 \varphi^4 + c_6 \varphi^6+ c_8 \varphi^8+\cdots 
While the Feynman rules for this theory are trivial, unlike those needed for the vectors and fermions in the DBIVA supermultiplet and the derivatively coupled pions in NLSM, the Feynman diagrams needed to compute amplitudes are identical to those of $A^{\text{OSS}}_{\times}$. Indeed, the two-loop amplitude for $\varphi^{2k}$ theory is generated simply by evaluating the following set of scalar 1-particle-reducible (1PI) integrals:

@@

Since the external states are massless, only the first two integrals are non-vanishing in dimensional regularization -- the rest are scaleless. Therefore, at least perturbatively, the above even-point theory of \eqn{eq:evenPointL} is equivalent to $\varphi^4$ theory at four-point through two-loop order: 
\begin{equation}
\mathcal{M}_{\varphi^{2k}}^{\text{2-loop}} \equiv \mathcal{M}_{\varphi^{4}}^{\text{2-loop}} 
\end{equation}
In addition to being perturbatively equivalent to the quartic sector of the theory at two-loop, we can see above that the integrals for the Feynman diagrams in \eqn{eq:2loopEven} are \textit{recursively one-loop}. In other words, to integrate the full amplitude at two-loop, we only need an iterated basis of one-loop integrals. This observation comes with the added advantage of permitting the application of one-loop Passarino-Veltman tensor reduction \cite{Passarino:1978jh} to two loop integrals. This property of recursively one-loop integrals will dramatically simplify the higher derivative EP theories that we consider in this work.

In what follows, we provide a catalog of the even-point effective field theories studied in this paper. We will define the Lagrangian, $\mathcal{L}$, to be the integrand of the action, $S$, when working in spacetime coordinate frame:
S= \int d^D x \,\mathcal{L}
Some theories in this work, like Volkov-Akulov and supersymmetric extensions of Born-Infeld, have natural formulations in terms of veirbeins that depend on fermionic degrees of freedom. For such theories, it is useful to distinguish between the coordinate independent actions, and the coordinate-dependent Lagrangians. 
\paragraph{Nonlinear Sigma Model}
The simplest effective field theory that we will consider in this work is the \textit{nonlinear sigma model} (NLSM). The NLSM Lagrangian can be expressed in terms of the chiral current, $j_\mu = U^\dagger \partial_\mu U$, as follows,
\be\label{eq:NLSMLag}
\mathcal{L}^{\text{NLSM}}= \frac{1}{2}\text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)] = \frac{1}{2}\text{tr}\left[\frac{(\partial_\mu\pi )( \partial^\mu \pi) }{(1-f_\pi^{-2}\pi^2)^2}\right]
where the trace is over the color indices of the gauge group, $\pi \equiv \pi^a T^a$, and on the right hand side we have applied the Cayley parameterization to the $SU(N_c)$ group elements, $U= \frac{1+\pi/f_\pi}{1-\pi/f_\pi}$, where $f_\pi$ is the dimensionful pion decay width. This theory is the unique two-derivative EFT that is invariant under constant shifts of the pion field in color space:
\be\label{eq:shiftSym}
\pi^a \rightarrow \pi^a + c^a
We could also consider higher derivative generalizations of the NLSM that are invariant under the shift symmetry of \eqn{eq:shiftSym}. Such EFTs are known formally as \textit{chiral perturbation theory}, and descriptors of pion scattering below the chiral symmetry breaking scale of QCD. 
\be\label{eq:XPTLag}
